---
title: "Data Exercise"
---


## Part 1 - Generating Synthetic Data


First, let's generate a synthetic dataset with 100 observations and 5 variables: age, sex, hypertension, cholesterol, and heart_health.


```{r}
#Load required packages
library(dplyr)
library(ggplot2)
library(gtsummary)

#Set seed for reproducibility
set.seed(123)

#Set sample size
n <- 100

#Create ages between 18 and 79
age <- sample(18:79, n, replace = TRUE)

#Create sexes (male or female)
sex <- sample(c("male", "female"), n, replace = TRUE)

#Create hypertension status (yes or no)
hypertension <- sample(c("yes", "no"), n, replace = TRUE)

#Create total cholesterol levels between 125 and 275, assuming males have higher total cholesterol than females
cholesterol <- ifelse(sex == "male", rnorm(n, mean = 220, sd = 20), rnorm(n, mean = 200, sd = 20))
cholesterol <- pmax(pmin(cholesterol, 275), 125)

#Create heart health scores between 25 and 90, assuming those with higher total cholesterol have higher heart health scores
heart_health <- 50 + 0.2 * cholesterol + rnorm(n, mean = 0, sd = 10)
heart_health <- pmax(pmin(heart_health, 90), 25)

#Create a data frame to store the generated data
heart <- data.frame(age = age, sex = sex, hypertension = hypertension, cholesterol = cholesterol, heart_health = heart_health)

#Get an overview and summary of the data
str(heart)
summary(heart)
```


## Part 2 - Processing Data


Then let's create two subset datasets, one with only males and one with only females.


```{r}
#Assign only males in the "heart" dataset to a new object
heart_males <- subset(heart, sex == "male")

#Assign only females in the "heart" dataset to a new object
heart_females <- subset(heart, sex == "female")

#Get an overview and summary of the new objects
str(heart_males)
summary(heart_males)

str(heart_females)
summary(heart_females)
```


Let's create a third subset dataset with only those with hypertension.


```{r}
#Create a new object by extracting only the data for those with hypertension from the "heart" dataset
heart_hypertension <- subset(heart, hypertension == "yes")

#Get an overview and summary of the new object
str(heart_hypertension)
summary(heart_hypertension)
```


## Part 3 - Exploring Data


Let's explore each variable in the dataset, starting with the outcome variable, heart_health.


```{r}
#Summary statistics and histogram for heart_health
heart$heart_health <- as.numeric(heart$heart_health)
summary(heart$heart_health)

ggplot(heart, aes(x = heart_health)) +
  geom_histogram() +
  labs(x = "Heart Health Score", y = "Frequency", title = "Histogram of Heart Health Score")
```


The mean (86) is less than the median (90) for the variable heart_health. The histogram of heart_health shows a left-skewing distribution with most observations at around x = 90.


Next let's explore the variable age.


```{r}
#Summary statistics and histogram for age
heart$age <- as.numeric(heart$age)
summary(heart$age)

ggplot(heart, aes(x = age)) +
  geom_histogram() +
  labs(x = "Age in Years", y = "Frequency", title = "Histogram of Age")
```


The mean (48) is slightly greater than the median (46) for the variable age. The histogram of age shows a relatively normal distribution.


Now let's explore the variable cholesterol.


```{r}
#Summary statistics and histogram for cholesterol
heart$cholesterol <- as.numeric(heart$cholesterol)
summary(heart$cholesterol)

ggplot(heart, aes(x = cholesterol)) +
  geom_histogram() +
  labs(x = "Cholesterol", y = "Frequency", title = "Histogram of Cholesterol")
```


The mean and median for the variable cholesterol are the same (209). The histogram of cholesterol shows a relatively normal distribution.


Next let's explore the variable sex.


```{r}
#Summary tables for the variable sex
table(heart$sex, useNA = "always")
prop.table(table(heart$sex))
```


55% of the observations are female, and 45% are male. There are no missing observations for sex.


Finally, let's explore the variable hypertension.


```{r}
#Summary tables for the variable hypertension
table(heart$hypertension, useNA = "always")
prop.table(table(heart$hypertension))
```


52% of the observations have hypertension, and 48% do not have hypertension. There are no missing observations for sex.


Here is a summary table of the outcome and predictors, stratified by sex.


```{r}
#Create a summary table of the variables, stratified by sex
heart %>% tbl_summary(by = sex, statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
  modify_caption("Summary Table of Heart Health Data, Stratified by Sex")
```


Here is another summary table of the outcome and predictors, stratified by hypertension.


```{r}
#Create a summary table of the variables, stratified by hypertension
heart %>% tbl_summary(by = hypertension, statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
  modify_caption("Summary Table of Heart Health Data, Stratified by Hypertension")
```


Now let's create some scatterplots between the main outcome (heart_health) and the continuous predictors, as well as a correlation table.


```{r}
#Create scatterplots between heart_health and the continuous predictors
pairs(cbind(heart_health, age, cholesterol))

#Create a correlation matrix of the continuous variables
cor(cbind(heart_health, age, cholesterol))
```


Based on the scatterplots and the correlation matrix, there doesn't seem to be a strong correlation between the outcome (heart_health) and any of the continuous predictors (age, cholesterol). 


Let's also look at the boxplots between the main outcome and the categorical variables.


```{r}
#Create a boxplot of sex and heart_health
ggplot(heart, aes(x = factor(sex), y = heart_health)) +
  geom_boxplot() +
  labs(x = "Sex", y = "Heart Health Score", title = "Boxplot of Sex and Heart Health Score")

#Create a boxplot of hypertension and heart_health
ggplot(heart, aes(x = factor(hypertension), y = heart_health)) +
  geom_boxplot() +
  labs(x = "Hypertension", y = "Heart Health Score", title = "Boxplot of Hypertension and Heart Health Score")
```


The boxplots of heart_health and the categorical variables (sex, hypertension) show negatively-skewed distributions.


Let's also look at the stratified boxplots.


```{r}
#Create a boxplot of sex and heart_health, stratified by hypertension
ggplot(heart, aes(x = factor(sex), y = heart_health, fill = factor(hypertension))) +
  geom_boxplot() +
  labs(x = "Sex", y = "Heart Health Score", title = "Boxplot of Sex and Heart Health Score, Stratified by Hypertension") +
  scale_fill_discrete(name = "Hypertension")

#Create a boxplot of hypertension and heart_health, stratified by sex
ggplot(heart, aes(x = factor(hypertension), y = heart_health, fill = factor(sex))) +
  geom_boxplot() +
  labs(x = "Hypertension", y = "Heart Health Score", title = "Boxplot of Hypertension and Heart Health Score, Stratified by Sex") +
  scale_fill_discrete(name = "Sex")
```


The stratified boxlplots show negatively-skewed distributions across all strata.


Next let's plot heart health as a function of age for males and females.


```{r}
#Plot heart health as a function of age for males and females
ggplot(heart_males, aes(x = age, y = heart_health)) + geom_point() + ggtitle("Age and Heart Health Score Among Males")

ggplot(heart_females, aes(x = age, y = heart_health)) + geom_point() + ggtitle("Age and Heart Health Score Among Females")
```


There is no noticeable correlation between age and heart health score for either males or females. There are several data points at heart health = 90, which is more than expected. This study sample may be healthier than the average population.


Then let's plot heart health as a function of cholesterol for males and females.


```{r}
#Plot heart health as a function of cholesterol for males and females
ggplot(heart_males, aes(x = cholesterol, y = heart_health)) + geom_point() + ggtitle("Total Cholesterol and Heart Health Score Among Males")

ggplot(heart_females, aes(x = cholesterol, y = heart_health)) + geom_point() + ggtitle("Total Cholesterol and Heart Health Score Among Females")
```


There is a positive correlation between total cholesterol and heart health score for males but no noticeable correlation or females. There are several data points at heart health = 90, which is more than expected. This study sample may be healthier than the average population.


Now let's plot heart health as a function of age for those with hypertension.


```{r}
#Plot heart health as a function of age for those with hypertension
ggplot(heart_hypertension, aes(x = age, y = heart_health)) + geom_point() + ggtitle("Age and Heart Health Score Among Those With Hypertension")
```


There is no noticeable correlation between age and heart health score for those with hypertension. There are several data points at heart health = 90, which is more than expected. This study sample may be healthier than the average population.


Finally, let's plot heart health as a function of cholesterol for those with hypertension.


```{r}
#Plot heart health as a function of cholesterol for those with hypertension
ggplot(heart_hypertension, aes(x = cholesterol, y = heart_health)) + geom_point() + ggtitle("Total Cholesterol and Heart Health Score Among Those With Hypertension")
```


There is a positive correlation between total cholesterol and heart health score for those with hypertension. There are several data points at heart health = 90, which is more than expected. This study sample may be healthier than the average population.


## Part 4 - Fitting Data


Let's fit a few simple models to the data. Let's start with fitting heart health as the outcome and age and cholesterol as the predictors for males.


```{r}
#Fit heart health as the outcome and age and cholesterol as the predictors for males
males <- lm(heart_health ~ age + cholesterol, data = heart_males)
summary(males)
```


Among males, total cholesterol is significantly associated with heart health score (p-value = 0.01), controlling for age; however, age is not significantly associated with hearth health score, controlling for total cholesterol (p-value = 0.64). The low R-squared value indicates that only 16% of the variation in heart health score is explained by the model.


Let's do the same model fitting for females.


```{r}
#Fit heart health as the outcome and age and cholesterol as the predictors for females
females <- lm(heart_health ~ age + cholesterol, data = heart_females)
summary(females)
```


Among females, neither age (p-value = 0.31) nor total cholesterol (p-value = 0.08) is significantly associated with heart health, controlling for other variables. The low R-squared value indicates that only 10% of the variation in heart health score is explained by the model.


Finally, let's fit heart health as the outcome and age and cholesterol as the predictors for those with hypertension.


```{r}
#Fit heart health as the outcome and age and cholesterol as the predictors for those with hypertension
htn <- lm(heart_health ~ age + cholesterol, data = heart_hypertension)
summary(htn)
```


Among those with hypertension, total cholesterol is significantly associated with heart health score (p-value = 0.01), controlling for age; however, age is not significantly associated with hearth health score, controlling for total cholesterol (p-value = 0.53). The low R-squared value indicates that only 13% of the variation in heart health score is explained by the model.

