---
title: "Data Exercise"
---

## Part 1 - Generating Synthetic Dataset and Checking Data

```{r}
library(dplyr)
library(ggplot2)

#Set seed for reproducibility
set.seed(123)

#Set sample size
n <- 100

#Create ages between 18 and 79
age <- sample(18:79, n, replace = TRUE)

#Create sexes (male or female)
sex <- sample(c("male", "female"), n, replace = TRUE)

#Create hypertension status (yes or no)
hypertension <- sample(c("yes", "no"), n, replace = TRUE)

#Create total cholesterol levels between 125 and 275, assuming males have higher total cholesterol than females
cholesterol <- ifelse(sex == "male", rnorm(n, mean = 220, sd = 20), rnorm(n, mean = 200, sd = 20))
cholesterol <- pmax(pmin(cholesterol, 275), 125)

#Create heart health scores between 25 and 90, assuming those with higher total cholesterol have higher heart health scores
heart_health <- 50 + 0.2 * cholesterol + rnorm(n, mean = 0, sd = 10)
heart_health <- pmax(pmin(heart_health, 90), 25)

#Create a data frame to store the generated data
heart <- data.frame(age = age, sex = sex, hypertension = hypertension, cholesterol = cholesterol, heart_health = heart_health)

#Get an overview and summary of the data
str(heart)
summary(heart)
```

## Part 2 - Processing Data

```{r}
#Assign only males in the "heart" dataset to a new object
heart_males <- subset(heart, sex == "male")

#Assign only females in the "heart" dataset to a new object
heart_females <- subset(heart, sex == "female")

#Get an overview and summary of the new objects
str(heart_males)
summary(heart_males)

str(heart_females)
summary(heart_females)
```

## Part 3 - Plotting Data

```{r}
#Plot heart health as a function of age for males and females
ggplot(heart_males, aes(x = age, y = heart_health)) + geom_point() + ggtitle("Age and Heart Health Score Among Males")

ggplot(heart_females, aes(x = age, y = heart_health)) + geom_point() + ggtitle("Age and Heart Health Score Among Females")
```

There is no noticeable correlation between age and heart health score for either males or females. There are several data points at heart health = 90, which is more than expected. This study sample may be healthier than the average population.

```{r}
#Plot heart health as a function of cholesterol for males and females
ggplot(heart_males, aes(x = cholesterol, y = heart_health)) + geom_point() + ggtitle("Total Cholesterol and Heart Health Score Among Males")

ggplot(heart_females, aes(x = cholesterol, y = heart_health)) + geom_point() + ggtitle("Total Cholesterol and Heart Health Score Among Females")
```

There is no noticeable correlation between total cholesterol and heart health score for either males or females. There are several data points at heart health = 90, which is more than expected. This study sample may be healthier than the average population.

## Part 4 - More Data Processing

```{r}
#Create a new object by extracting only the data for those with hypertension from the "heart" dataset
heart_hypertension <- subset(heart, hypertension == "yes")

#Get an overview and summary of the new object
str(heart_hypertension)
summary(heart_hypertension)
```

## Part 5 - More Data Plotting

```{r}
#Plot heart health as a function of age for those with hypertension
ggplot(heart_hypertension, aes(x = age, y = heart_health)) + geom_point() + ggtitle("Age and Heart Health Score Among Those With Hypertension")
```

There is no noticeable correlation between age and heart health score for those with hypertension. There are several data points at heart health = 90, which is more than expected. This study sample may be healthier than the average population.

```{r}
#Plot heart health as a function of cholesterol for those with hypertension
ggplot(heart_hypertension, aes(x = cholesterol, y = heart_health)) + geom_point() + ggtitle("Total Cholesterol and Heart Health Score Among Those With Hypertension")
```

There is no noticeable correlation between total cholesterol and heart health score for those with hypertension. There are several data points at heart health = 90, which is more than expected. This study sample may be healthier than the average population.

## Part 6 - Fitting Simple Models

```{r}
#Fit heart health as the outcome and age and cholesterol as the predictors for males
males <- lm(heart_health ~ age + cholesterol, data = heart_males)
summary(males)
```

Among males, total cholesterol is significantly associated with heart health score (p-value = 0.01), controlling for age; however, age is not significantly associated with hearth health score, controlling for total cholesterol (p-value = 0.64). The low R-squared value indicates that only 16% of the variation in heart health score is explained by the model.

```{r}
#Fit heart health as the outcome and age and cholesterol as the predictors for females
females <- lm(heart_health ~ age + cholesterol, data = heart_females)
summary(females)
```

Among females, neither age (p-value = 0.31) nor total cholesterol (p-value = 0.08) is significantly associated with heart health, controlling for other variables. The low R-squared value indicates that only 10% of the variation in heart health score is explained by the model.

```{r}
#Fit heart health as the outcome and age and cholesterol as the predictors for those with hypertension
htn <- lm(heart_health ~ age + cholesterol, data = heart_hypertension)
summary(htn)
```

Among those with hypertension, total cholesterol is significantly associated with heart health score (p-value = 0.01), controlling for age; however, age is not significantly associated with hearth health score, controlling for total cholesterol (p-value = 0.53). The low R-squared value indicates that only 13% of the variation in heart health score is explained by the model.
